---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r }

knitr::opts_chunk$set(echo = TRUE)

```




##Load libraries
```{r }
library(ggplot2)
library(dplyr)
```

## Loading and preprocessing the data
```{r }
my_data<-read.csv("~/MyProjects/representative projects/week2/activity.csv", header=TRUE)
head(my_data)
tail(my_data)
str(my_data)

```


## What is mean total number of steps taken per day?
##2.histogram of number steps taken each day
#Summarise data, removing missing numbers, formating dates and summing up total number of steps taken
```{r}
my_data$newformat<-as.character(my_data$date)
my_data$newformat<-as.Date(my_data$newformat)

k<-my_data%>%
        select(steps, newformat)%>%
        filter(!is.na(steps))%>%
        group_by(newformat)%>%
        summarise(SumSteps=sum(steps))

```

#Putting the numbers on histogram

```{r}
hist(k$SumSteps, main = "Distribution of Total Number of Steps", 
     xlab = "Number of Steps", 
     ylab = "Frequency", 
     breaks=10,   col = "navy")

```

#3.Mean and median number of steps taken each day
```{r}
summary(k$SumSteps)
```




## What is the average daily activity pattern?

#4. 5 minutes time series plot of the average number of steps taken
# Takinf the means of each 5 minutes interval, exluding missing data
```{r}

j<-my_data%>%
        select(steps, interval)%>%
        filter(!is.na(steps))%>%
        group_by(interval)%>%
        summarise(AvgSteps=mean(steps))
j

```

#Plotting the data, formating the intervals into the real time of he day
```{r}

ggplot(j, aes(y=AvgSteps, x=strptime(sprintf("%04d", j$interval), format="%H%M")))+geom_line()+labs(x= "Time of the day") + labs(y="Average Steps taken")+labs(title="Pattern of a day")


```

Here we see the patterns of the day

#5.The 5-minute interval that, on average, contains the maximum number of steps
```{r}

   j%>%
        select(AvgSteps, interval)%>%
         filter(AvgSteps==max(AvgSteps))

      

```

Here we see that the average maximum number of steps was taken in the 835th inteval



## Imputing missing values

#6.Code to describe and show a strategy for imputing missing data
```{r}
n.missing<-sum(is.na(my_data))#number of missing values in the dataset

```


Imputting missing data by mean number of steps taken a day.
```{r}
 

StepsImputed<- inner_join(my_data, j,  by="interval") %>% 
                    mutate(steps=ifelse(is.na(steps), AvgSteps, steps)) %>%
                    select(date, interval, steps)
head(StepsImputed)

```


## Are there differences in activity patterns between weekdays and weekends?

Summary of the new data, wee see there is no NAs

```{r}
summary(StepsImputed)
```


it's time to see new averages of steps done per day
```{r}
z<-StepsImputed%>%
        select(steps, date)%>%
        group_by(date)%>%
        summarise(SumSteps=sum(steps))
z 

summary(z$SumSteps)
```

Wee see that the median increased by 10 steps/day.


 #It's time to plot

```{r}

hist(z$SumSteps,  main = "New Distibution of Steps per day", 
     xlab = "Steps per Day", 
     ylab = "Frequency ", 
     breaks=20,   col = "grey")

```

#8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends


Determine the weekdays and weekends, based on date column, with imputed data.
```{r}
StepsImputed$date<-as.Date(StepsImputed$date)
StepsImputed <- StepsImputed %>%
    mutate(Weekdays = ifelse(weekdays(date) %in% c("Saturday","Sunday"),
                                     "weekend","weekday"))
           
head(StepsImputed)
```

Number of steps taken by the day of the week
```{r}
a<-StepsImputed%>%
        select(steps, interval, Weekdays)%>%
        group_by(Weekdays, interval )%>%
        summarise(AvgSteps=mean(steps))
a
```


```{r}
ggplot(a, aes(y=AvgSteps, x=strptime(sprintf("%04d", a$interval), format="%H%M")))+geom_line(aes(col=Weekdays))+labs(x= "Time of the day") + labs(y="Average Steps taken")

```



